Election 2024

### 2024-10-23 paper/paper.qmd
update paper.qmd with discussion section

The logistic regression model provides a structured approach to predict the likelihood of Trump winning in various states, based on recent and high-quality polling data. By focusing on key features such as polling percentages and sample sizes, the model captures the competitive dynamics between candidates and offers valuable insights into potential election outcomes. Future iterations of this model can incorporate real-time data updates to enhance predictive accuracy as the election approaches.
 # 5 Results {#sec-results}
 # 6 Discussion {#sec-discussion}

>>>>

The logistic regression model provides a structured approach to predict the likelihood of Trump winning in various states, based on recent and high-quality polling data. By focusing on key features such as polling percentages and sample sizes, the model captures the competitive dynamics between candidates and offers valuable insights into potential election outcomes. Future iterations of this model can incorporate real-time data updates to enhance predictive accuracy as the election approaches.

 # 4 Results {#sec-results}
The results of our analysis are as follows. We have Donald Trump predicted to win Arizona, Georgia and Michigan for a total of 42 of the 93 swing state votes. While Kamala Harris is projected to win Nevada, North Carolina, Pennsylvania and Wisconsin for the remaining 51 swing state votes. This results in a win for the democrats and the first female US president!

# 5 Discussion {#sec-discussion}

## What Is Done in This Paper?

In this paper, we create a logistic regression model to predict the outcome of the 2024 U.S. Presidential Election, focusing on state-level polling data to estimate the probability of Donald Trump winning against Kamala Harris. By aggregating recent and reliable polls for swing states and weighting them by sample size and recency, we aim to provide a data-driven analysis of voter preferences. Our model is constructed to identify the likelihood of Trump securing a majority in each state, ultimately offering a prediction for the overall election result.

## What Do We Learn About the World?

One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets. While the majority of reputable polls suggest that Kamala Harris is favored on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket]. This stark disconnect raises questions about the accuracy of traditional polling methods and whether betting markets, which are financial tools driven by market forces, may offer a more precise reflection of public sentiment.

At first glance, one might expect that if polling data were more accurate than betting markets, arbitrage opportunities would emerge, allowing savvy participants to profit from discrepancies. However, this does not seem to be happening, which suggests that the markets may be pricing in information that the polls do not capture—perhaps reflecting shifts in voter sentiment, hidden preferences, or systematic biases in polling.

## What Else Do We Learn About the World?

Another important consideration is the possible biases in both polling and betting markets. Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favor of Harris. Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favor him. Furthermore, the rapid expansion and increased liquidity of betting markets in the last few years may have improved their efficiency, making them more reflective of real-world probabilities. However, historical data still suggest that polls have been accurate 78% of the time in predicting elections [@FiveThirtyEight], indicating that the reliability of betting markets remains questionable, especially in light of their poor performance during the 2016 election, when Trump’s odds were listed at +475 (17%) just before his victory [https://www.oddsshark.com/entertainment/us-presidential-odds-2016-futures].

## Weaknesses of the Model

While our model provides a structured framework for predicting the election, it has several limitations. First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals). Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve non-linear influences that are difficult to predict.

Moreover, our focus on swing states limits the model’s applicability to the national picture. While swing states are crucial to the election outcome, non-swing states could offer additional insights into broader voter trends, and including them in future analyses could enhance the model’s robustness. Training the model on a more comprehensive dataset that includes these states and other predictive variables, such as demographic factors or economic indicators, could lead to a more accurate forecast.

## How Should We Proceed in the Future?

Looking ahead, future iterations of this model should incorporate more diverse data sources. Expanding the dataset to include polling data from all states, along with betting market information, could improve prediction accuracy. Moreover, experimenting with different types of predictive models, such as Bayesian logistic regression or machine learning models like random forests, could offer more sophisticated insights and account for non-linear interactions that our current model might miss.

It would also be valuable to study the interaction between polling data and betting markets more closely, potentially integrating them into a unified prediction model. This hybrid approach might help reconcile the discrepancies observed between the two sources and provide a more nuanced understanding of election dynamics. Additionally, out-of-sample validation and sensitivity analyses should be conducted to test the robustness of our model and adjust for overfitting.

Ultimately, while our model offers a strong foundation for predicting the 2024 U.S. Presidential Election, there is still much to learn. By refining the model and incorporating additional data, we can move closer to producing predictions that more accurately reflect voter behavior and election outcomes.
